---
title: "assignment1"
author: "Hyunjung Byun"
date: "2018<eb>7<ec>13<ec>"
output:
html_document:
df_print: paged
---
<br><br>
<h3>QUESTION 1<br></h3>

central tendency: 
mode, median, mean
<br>
variation:
range, qualtile deviation, variance, standard deviation
<br><br>

<h3>QUESTION 2<br></h3>
first way: using <strong>c()</strong> function which combines vectors<br>
second way : using <strong>seq()</strong> function which  generates a sequence of numbers<br>
third way : using <strong>rep()</strong> function which rep() function replicates the vectors<br><br>


<h3>QUESTION 3<br></h3>
answer: character<br>
```{r}
x3<-c('x','x','x',1,3,5,7,9,2,4,6,8,10)
class(x3)
```
<br><br>
<h3>QUESTION 4<br></h3>

```{r}
x4<-seq(1,100,by=2)
x4
```

<br><br>
<h3>QUESTION 5<br></h3>
```{r}
x5<-x4[x4 <=60 | x4 >=80]
x5
```

<br><br>
<h3>QUESTION 6<br></h3>
```{r}
ff <- function(x){
  mean = mean(x)
  median = median(x)
  sd = sd(x)
  return(c(mean, median, sd))
}
ff(x5)
```

<br><br>
<h3>QUESTION 7</br></h3>

<strong>first method </strong>: numbers by column in 3ways<br>
belows all have the same result<br>
```{r}
x1<-c(2,3,7,1,6,2,3,5,1)
mat1<-matrix(x1,ncol=3) 
#or
mat1<-matrix(x1,3,3)
#or
mat1<- cbind(c(2,3,7),c(1,6,2),c(3,5,1))
mat1
```
<strong>second method</strong>: numbers by row<br>
```{r}
x2<-c(3,2,9,0,7,8,5,8,2)
mat2<-matrix(x2,ncol=3, byrow=TRUE)
mat2
```

<br><br>
<h3>QUESTION 8<br></h3>
```{r}

matriceproduct<-mat1 %*% mat2
matriceproduct

```
<h3>QUESTION 9<br></h3>
<strong>*Find the class of ‘iris’ dataframe<br></strong>
```{r}
data.frame(iris)
```
<br>
<strong>*find the class of all the columns of ‘iris' get the summary</strong>
```{r}
lapply(iris, class) 
#or
sapply(iris, class)
```
<br>
<strong>*Get rownames, get column names</strong>
```{r}
rownames(iris)
colnames(iris)
```
<br>
<strong>*Get the number of rows and number of columns.</strong><br>
answer: rows-> 150, columns -> 5
```{r}
str(iris)
#or
nrow(iris)
ncol(iris)
```
<br><br>

<h3>QUESTION 10</h3>
```{r}
x10<-iris[(nrow(iris)-1):nrow(iris),(ncol(iris)-1):ncol(iris)]
x10
```
