---
title: "Assignment3"
author: "Hyunjung Byun"
---


<h3>Q1. Import the data into R.</h3>
```{r}
library(dplyr)
forestfires = read.csv("c:/users/byunh/forestfires.csv", header = TRUE, sep = ',')
forestfires
```

<h4>a. How many observations are there in the dataset?</h4>
```{r}
nrow(forestfires)

#or

forestfires %>% summarize(observations=n())
```

<h4>b. How many observations are there with a fire (i.e.,area>0) </h4>
```{r}
nrow(forestfires[forestfires$area>0, ])

#or

nrow(subset(forestfires, area>0))
```

<h4>c. How many observations are there with a rain(i.e., rain>0)</h4>
```{r}
nrow(forestfires[forestfires$rain>0, ])

#or

nrow(subset(forestfires, rain>0))
```

<h4>d. How many observations are there with both a fire and a rain? </h4>
```{r}
nrow(forestfires[forestfires$area>0 & forestfires$rain>0, ])

#or

forestfires %>% filter(area>0 & rain>0) %>% summarise(observations=n())
```

<h3>Q2. Show the columns month, day, area of the all the observations. </h3>
```{r}
forestfires[c('month', 'day','area')]

#or

forestfires %>% select(month, day, area)
```

<h3>Q3. Show the columns month, day, area of the observations with a fire.</h3>
```{r}
forestfires %>% filter(area>0) %>% select(month, day, area)
```

<h3>Q4. How large are the five largest fires (i.e., having largest area) </h3>
```{r}
forestfires %>% arrange(desc(area)) %>% top_n(5) %>% select(area)
```

<h4>a. What are the corresponding month, temp, RH, wind, rain area? </h4>
```{r}
forestfires %>% arrange(desc(area)) %>% top_n(5) %>% select(month, temp, RH, wind, rain, area)
```

<h4>b. Add one column to the data indicating whether a fire occurred for each observation(True for area >0 and False for area ==0) </h4>
```{r}
forestfires$fire = factor(forestfires$area>0)

#OR

forestfires = forestfires %>% mutate(fire=area>0)
forestfires
str(forestfires)
```

<h3>Q5. Construct a matrix with 10 columns and 10 rows,filled with random numbers
between 0 and 1. Apply two different methods.</h3>
```{r}
#First way
mat<-matrix(runif(100), ncol=10)
mat
```
```{r}
#Sencond way
mat<-matrix(data=runif(100), nrow=10, ncol=10, byrow =TRUE)
mat
```

<h4>a. Calculate the row means of this matrix. </h4>
```{r}
rowMeans(mat)
```

<h4>b. Remake the above matrix, with 100 columns, and 10 rows. Then calculate the column means.</h4>
```{r}
mat2<-matrix(runif(1000), ncol=100)
mat2
```
```{r}
colMeans(mat2)
```

