---
title: "Assignment4"
author: "Hyunjung Byun"
---
---
title: "Assignment4"
author: "Hyunjung Byun"

---

#Load	 worldrecord.csv	 and	 answer	 the	following?
```{r}
library(dplyr)
WorldRecords = read.csv("c:/users/byunh/WorldRecords.csv", header = TRUE, sep = ',')
WorldRecords
```

#1. How	 many	 different	 types	 of	 events	 (e.g.	
#“Mens	 100m”,	 “Womens	 shotput”	 etc)	 are	represented	in	the	dataset
```{r}
length(unique(WorldRecords$Event))
```

# 2. In what year did Usain Bolt first break the world record for the men's 100m?
```{r}
A<-WorldRecords[WorldRecords$Athlete == 'Usain Bolt' ,]
A[1,]$Year
```

#3. Which	variable	tells	us	the	record	setting	time	or	distance?
# type
```{r}
str(WorldRecords)
unique(WorldRecords$Type)
print("Type")
```

#The	variable	name	in	the	data	set	is?	
```{r}
str(WorldRecords)
print("Type")
```

#What	type	of	the	variable	is	this?
```{r}
class(WorldRecords$Type)
```

#4. Create	 a	 subset	 of	 the	 dataset	 that	 contains	
#only	the	world	record	cases	for	men’s	shotput	and	women’s	shotput
```{r}
Shotput<-subset(WorldRecords, Event %in% c('Mens Shotput','Womens Shotput'))
Shotput
```

#5 .Create	 a	 scatter	 plot	 of	 the	 year	 and	 record	shotput	 distance	 one	 for	 #men	 and	 one	 for	women.
```{r}
library(ggplot2)
mensShotput<- filter(Shotput, Event=="Mens Shotput")
ggplot(mensShotput, aes(x = Year, y = Record))+ geom_point() + labs(title="Mens Shotput")

womensShotput<- filter(Shotput, Event=="Womens Shotput")
ggplot(womensShotput, aes(x = Year, y = Record))+ geom_point() + labs(title="Womens Shotput")
```


#6. Find	 the	 average/mean	 	 time	 for	 each	 event.	
```{r}
X<- filter(WorldRecords, Type=="time")
Y<- group_by(X, Event)
summarise(Y, AvgTime=mean(Record))
```

#How	 many	 athletes	 have	 time	 more	 than	average	in	each	event.
```{r}
Z<- mutate(Y, AvgTime=mean(Record))
W<- filter(Z, Record > AvgTime)
summarise(W, Athletes=n())
```

#7. Select	the	athlete	who	took	most	time	in	men’s	s100m		and	women’s	event.
#If you ask who took the longest time, The answer should be
```{r}
A<- filter(WorldRecords, Event=="Mens 100m")
B<- filter(A, Record==max(Record))
C<-select(B, Athlete, Event, Record)
C

A<- filter(WorldRecords, Event=="Womens 100m")
B<- filter(A, Record==max(Record))
C<-select(B, Athlete, Event, Record)
C
```
#but if you ask who have the most participations,The answer should be
```{r}
A<- filter(WorldRecords, Event=="Mens 100m")
B <- group_by(A, Athlete)
C<- summarize(B, NumberOfRecords=n())
filter(C, NumberOfRecords == max(NumberOfRecords))

A<- filter(WorldRecords, Event=="Womens 100m")
B<- group_by(A, Athlete)
C<- summarize(B, NumberOfRecords=n())
filter(C, NumberOfRecords == max(NumberOfRecords))
```

#8. Which	 country	won	maximum	 times	 of	men’s	100m	event?
```{r}
A <- filter(WorldRecords, Event=="Mens 100m")
B <- group_by(A, Nationality)
C <- summarize(B, NumberOfRecords=n())
filter(C, NumberOfRecords == max(NumberOfRecords))
```

#9. How	many	athletes	are	there	in	each	event?
```{r}
Y<-distinct(WorldRecords,Event, Athlete)
aggregate(Athlete ~ Event, data = Y, FUN = length)
```

#10. Which	country	has	maximum	wins?
```{r}
unique(WorldRecords$Nationality)

#unify some countries' name
WorldRecords$Nationality[WorldRecords$Nationality=="?United States"] <- "United States"
WorldRecords$Nationality[WorldRecords$Nationality=="USA"] <- "United States"
WorldRecords$Nationality[WorldRecords$Nationality=="?East Germany"] <- "GER"
WorldRecords$Nationality[WorldRecords$Nationality=="?West Germany"] <- "GER"
WorldRecords$Nationality[WorldRecords$Nationality=="GDR"] <- "GER"
WorldRecords$Nationality[WorldRecords$Nationality=="FRG"] <- "GER"
WorldRecords$Nationality[WorldRecords$Nationality=="?Italy"] <- "ITA"
WorldRecords$Nationality[WorldRecords$Nationality=="?New Zealand"] <- "NZL"
WorldRecords$Nationality[WorldRecords$Nationality=="?United Kingdom"] <- "GBR"
WorldRecords$Nationality[WorldRecords$Nationality=="?Australia"] <- "AUS"
WorldRecords$Nationality[WorldRecords$Nationality=="?France"] <- "FRA"
WorldRecords$Nationality[WorldRecords$Nationality=="?Sweden"] <- "SWE"
WorldRecords$Nationality[WorldRecords$Nationality=="?Soviet Union"] <- "?Russia"
WorldRecords$Nationality[WorldRecords$Nationality=="?Poland"] <- "POL"
WorldRecords$Nationality[WorldRecords$Nationality=="URS"] <- "?Russia"

A<- group_by(WorldRecords, Nationality)
B <- summarize(A, NumberOfRecords=n())
filter(B, NumberOfRecords == max(NumberOfRecords))
```
